---
title: "p8105_hw1_yx2638"
author: "Yifei Xu"
date: "2022-09-20"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

## Load the penguins dataset
```{r}
data("penguins", package = "palmerpenguins")
```

## Write a short description of the penguins dataset 
```{r}
# names / values of important variables
variable_name = names(penguins)
species_name = levels(penguins$species)
island_name = levels(penguins$island)
sex_name = levels(penguins$sex)

## number of each species
n_adelie = sum(pull(penguins, species) == "Adelie")
n_gentoo = sum(pull(penguins, species) == "Gentoo")
n_hinstrap = sum(pull(penguins, species) == "Chinstrap")

## bill length
mean_bill_length = penguins %>%
        drop_na() %>%
        pull(bill_length_mm) %>%
        mean() %>%
        round(digits = 1)

max_bill_length = penguins %>%
        drop_na() %>%
        pull(bill_length_mm) %>%
        max() 

min_bill_length = penguins %>%
        drop_na() %>%
        pull(bill_length_mm) %>%
        min() 
        
 ## bill depth
mean_bill_depth = penguins %>%
        drop_na() %>%
        pull(bill_depth_mm) %>%
        mean() %>%
        round(digits = 1)

max_bill_depth = penguins %>%
        drop_na() %>%
        pull(bill_depth_mm) %>%
        max() 

min_bill_depth = penguins %>%
        drop_na() %>%
        pull(bill_depth_mm) %>%
        min() 
        
 ## flipper length
mean_flipper_length = penguins %>%
        drop_na() %>%
        pull(flipper_length_mm) %>%
        mean() %>%
        round(digits = 0)

max_flipper_length = penguins %>%
        drop_na() %>%
        pull(flipper_length_mm) %>%
        max() 

min_flipper_length = penguins %>%
        drop_na() %>%
        pull(flipper_length_mm) %>%
        min() 

## body mass
mean_body_mass = penguins %>%
        drop_na() %>%
        pull(body_mass_g) %>%
        mean() %>%
        round(digits = 0)

max_body_mass = penguins %>%
        drop_na() %>%
        pull(body_mass_g) %>%
        max() 

min_body_mass = penguins %>%
        drop_na() %>%
        pull(body_mass_g) %>%
        min() 
        
 
# size of the dataset 
row = nrow(penguins)
col = ncol(penguins)

# mean flipper length
mean_flipper_length = penguins %>%
        drop_na() %>%
        pull(flipper_length_mm) %>%
        mean() %>%
        round(digits = 2)
```
There are `r col` variables in this dataset: `r variable_name`.

The levels of species are `r species_name`, the levels of island are `r island_name` and the levels of sex are `r sex_name`.

For the species, there are `r n_adelie` Adelie penguins, `r n_gentoo` Gentoo penguins and `r n_hinstrap` H Instrap penguins in this dataset.

The mean of bill length is `r mean_bill_length` mm, the maximum is `r max_bill_length` mm and the minimum is `r min_bill_length` mm.

The mean of bill depth is `r mean_bill_depth` mm, the maximum is `r max_bill_depth` mm and the minimum is `r min_bill_depth` mm.

The mean of flipper length is `r mean_flipper_length` mm, the maximum is `r max_flipper_length` mm and the minimum is `r min_flipper_length` mm.

The mean of body mass is `r mean_body_mass` g, the maximum is `r max_body_mass` g and the minimum is `r min_body_mass` g.

Totally, there are `r row` rows and `r col` columns in the dataset. 


## Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x)
```{r}
# draw the plot
penguine_plot = penguins %>%
        drop_na() %>%
        ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +  geom_point(aes(color = species))

# save the plot
ggsave("hw1_scatterplot.pdf", plot = penguine_plot)

```

