---
title: "p8105_hw1_yx2638"
author: "Yifei Xu"
date: "2022-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
```

# Problem 1

### Load the penguins dataset
```{r}
data("penguins", package = "palmerpenguins")
```

### Write a short description of the penguins dataset 
```{r, message=FALSE}
# summary of the dataset
skimr::skim(penguins)

# names / values of important variables
variable_name = names(penguins)

species_name = levels(penguins$species)
island_name = levels(penguins$island)
sex_name = levels(penguins$sex)

# size of the dataset 
row = nrow(penguins)
col = ncol(penguins)

# mean flipper length
mean_flipper_length = penguins %>%
        drop_na() %>%
        pull(flipper_length_mm) %>%
        mean() %>%
        round(digits = 2)
```
There are `r col` variables in this dataset: `r variable_name`.<br> 
The levels of species are `r species_name`, the levels of island are `r island_name` and the levels of sex are `r sex_name`.<br> 
As for the size of the dataset, there are `r row` rows and `r col` columns in total. <br> 
The mean of flipper length is `r mean_flipper_length` mm. <br>


### Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x)
```{r, message=FALSE}
# draw the plot
penguine_plot = penguins %>%
        drop_na() %>%
        ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +  geom_point(aes(color = species))

# save the plot
ggsave("hw1_scatterplot.pdf", plot = penguine_plot)

```


# Problem 2

### Create a dataframe
```{r, message=FALSE}
df = tibble (
        vec_norm = rnorm(n = 10),
        vec_log = vec_norm > 0,
        vec_char = c("a", "b", "c", "d","e","f","g","h","i","j"),
        vec_fac = factor(c("High", "Low", "Medium", "High", "Low", "Low", "Medium", "High", "Low", "Medium"))
)
```

### Take the mean of each variable in the dataframe
```{r, message=FALSE}
# mean of the random sample from standard normal distribution
mean_norm = df %>%
        pull(vec_norm) %>%
        mean()
mean_norm

# mean of logical vector
mean_log = df %>%
        pull(vec_log) %>%
        mean()
mean_log

# mean of character vector
mean_char = df %>%
        pull(vec_char) %>%
        mean()
mean_char

# mean of factor vector
mean_fac = df %>%
        pull(vec_fac) %>%
        mean()
mean_fac
```
The `mean` function works for the numeric and logical vectors while fail to work for the character and factor vector.

### Convert variables to numeric
```{r, message=FALSE}
# convert logical variable to numeric and calculate the mean
mean_log_num = df %>%
        pull(vec_log) %>%
        as.numeric() %>%
        mean()
mean_log_num

# convert character variable to numeric and calculate the mean
mean_char_num = df %>%
        pull(vec_char) %>%
        as.numeric() %>%
        mean()
mean_char_num

# convert factor variable to numeric and calculate the mean
mean_fac_num = df %>%
        pull(vec_fac) %>%
        as.numeric() %>%
        mean()
mean_fac_num
```
After converting to numeric, the `mean` function can work for logical and factor vector. However, it still fails to work for character vector. This is because the `as.numeric` function sets TRUE to 1 and FALSE to 0 for the logical vector and sets "High" to 1, "Low" to 2, "Medium" to 3 for the factor vector. All values in character vector return NA.

